State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-04-28 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
its just a check
paper updates

%<-comment out this for checking equations (author list too)
\usepackage{iopams,setstack} 
%<-comment out this for checking equations
\usepackage{graphicx,epsfig}
\usepackage{algorithm, algorithmicx, algpseudocode} 
\usepackage{color} 
\usepackage[tight,scriptsize]{subfigure}
\usepackage{pstool}
%\renewcommand{\subcapsize}{\small}
% \usepackage{anysize}							%<-use this for checking equation length
% \marginsize{1.7cm}{1.5cm}{0.3cm}{0.8cm}			%<-use this for checking equation length
% \setlength{\columnsep}{0.5cm}					%<-use this for checking equation length
% \usepackage{amsmath,amssymb,amsfonts} 			%<-use this for checking equation length
\newcommand{\todo}[1]{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}} 
\newcommand{\wtf}[1]{\textsf{\emph{\textbf{\textcolor{red}{#1}}}}} 
\newcommand{\omg}[1]{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}} \begin{document}

@@

The work from Schiff and Sauer \cite{schiff2008kalman} successfully demonstrated it is possible to estimate parameters \wtf{what kind of parameters?} from the WC equations. This motivates the question, what parameters are the most important? How patient specific do the assumed parameters have to be for the model to be useful? In this paper, we address both of these questions and provide a theoretical platform to perform patient-specific grey-box modelling of the human neocortex. \wtf{not sure about rhetorical questions as a narrative mechanism in general, also not sure how this paper addresses these particular questions.} \omg{Maybe the next section addresses that the connectivity is important. Needs sharpening.}

@@

The firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function 
\begin{equation}
	\label{ActivationFunction} f\left( v\left( \mathbf{r}', t \right) \right) = \frac{f_{max}\left(\mathbf{r}\right)}{1 + \exp \left( \varsigma \left( v_0\left( \mathbf{r} \right) - v\left(\mathbf{r}',t\right) \right) \right)}. 
\end{equation}

%Even though we will model $v_0$ and $\nu_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality at this point. 
By substituting \ref{RateBasedInteractions} into \ref{SpikesToPotential} we get the spatiotemporal model 
\begin{equation}
	\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) = \int_{-\infty}^t h\left(t - t'\right) \left(\int_\Omega w\left(\mathbf{r},\mathbf{r}'\right) f\left( v\left( \mathbf{r}',t \right)\right)d\mathbf{r}' + u(\mathbf{r},t) \right)dt'. 
\end{equation}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function 
\begin{equation}
	\label{GreensFuncDef} Dh\left( t \right) = \delta \left( t \right), 
\end{equation}
where $D=\frac{d}{dt} + \zeta$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{equation}
	\label{FinalFormContinuous} \frac{dv\left( \mathbf{r},t \right)}{dt} + \zeta v\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)dr'} + u\left(\mathbf{r},t\right). 
\end{equation}
To arrive at the integro-difference equation (IDE) form of the model we discretize time using a first-order Euler method (see~\ref{Time Discretization}) giving 
\begin{equation}
	\label{DiscreteTimeModel} v_{t+1}\left(\mathbf{r}\right) = \xi v_t\left(\mathbf{r}\right) + T_s \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} + T_s u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right), 
\end{equation}
where $T_s$ is the time step, $\xi = 1-\zeta T_s$ and $e_t\left(\mathbf{r}\right)$ is an iid spatially coloured, temporally white disturbance term with the covariance 
\begin{equation}
	cov(\mathbf{e}_t,\mathbf{e}_{t+\tau}) = \left\{ 
	\begin{array}{*{20}
		{c}} \gamma\left(\mathbf{r}-\mathbf{r'}\right), & \tau=0 \\
		0 & \mathrm{otherwise}. \\
	\end{array}
	\right. 
\end{equation}
This term was added to drive the system and account for uncertainty in the model. To generate iEEG/LFP data we used the output function 
\begin{equation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)d\mathbf{r}'} + \boldsymbol{\varepsilon}_t, 
\end{equation}
where $\mathbf{r}_n$ defines the location of the sensors in the field where $n=1,...,N$ indexes the sensors and $\boldsymbol{\varepsilon}_t \sim \mathcal{N}\left(0,\Sigma{\varepsilon}\right)$. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the geometry of pick-up range of sensors where 
\begin{equation}
	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)}. 
\end{equation}

@@

	\label{reduced continuous model} \boldsymbol{\phi}^{\top}\left( \mathbf{r} \right) \mathbf{x}_{t+1} = T_s\boldsymbol{\theta}^{\top}\int_\Omega {\boldsymbol{\psi} \left(\mathbf{r}-\mathbf{r}'\right) f\left( \boldsymbol{\phi} ^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t \right)d\mathbf{r}'} - \xi\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right)x_t + T_s u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right). 

@@

	\label{Homogeneous SS Model} \mathbf{x}_t &=& q(\mathbf{x}_t)\boldsymbol{\theta} - \xi\mathbf{x}_t + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	&=& Q\left(\mathbf{x}_t,u_t(\mathbf{r})\right) + \mathbf{e}_t \\

@@

\psfragfig{./Graph/FreqAnalysis}

>>>>

%<-comment out this for checking equations (author list too)
\usepackage{iopams,setstack} 
%<-comment out this for checking equations
\usepackage{graphicx,epsfig}
\usepackage{algorithm, algorithmicx, algpseudocode} 
\usepackage{color} 
\usepackage[tight,scriptsize]{subfigure}
% \usepackage{pstool}
\renewcommand{\subcapsize}{\small}
% \usepackage{anysize}							%<-use this for checking equation length
% \marginsize{1.7cm}{1.5cm}{0.3cm}{0.8cm}			%<-use this for checking equation length
% \setlength{\columnsep}{0.5cm}					%<-use this for checking equation length
\usepackage{amssymb,amsfonts} 			%<-use this for checking equation length
\newcommand{\todo}[1]{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}} 
\newcommand{\wtf}[1]{\textsf{\emph{\textbf{\textcolor{red}{#1}}}}} 
\newcommand{\omg}[1]{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}} \begin{document}

@@

The work from Schiff and Sauer \cite{schiff2008kalman} successfully demonstrated it is possible to perform state estimation of modified WC equations. This motivates the question, what parameters are the most important? How patient specific do the assumed parameters have to be for the model to be useful? In this paper, we address both of these questions and provide a theoretical platform to perform patient-specific grey-box modelling of the human neocortex. \wtf{not sure about rhetorical questions as a narrative mechanism in general, also not sure how this paper addresses these particular questions.} \omg{Maybe the next section addresses that the connectivity is important. Needs sharpening.}

@@

The firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function 
\begin{equation}
	\label{ActivationFunction} f\left( v\left( \mathbf{r}', t \right) \right) = \frac{f_{max}\left(\mathbf{r}\right)}{1 + \exp \left( \varsigma \left( v_0\left( \mathbf{r} \right) - v\left(\mathbf{r}',t\right) \right) \right)}. 
\end{equation}
%Even though we will model $v_0$ and $\nu_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality at this point. 
By substituting \ref{RateBasedInteractions} into \ref{SpikesToPotential} we get the spatiotemporal model 
\begin{eqnarray}
	\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) =  \int_{-\infty}^t h\left(t - t'\right) \\
	\times\left(\int_\Omega w\left(\mathbf{r},\mathbf{r}'\right) f\left( v\left( \mathbf{r}',t \right)\right)d\mathbf{r}' + u(\mathbf{r},t) \right)dt'. 
\end{eqnarray}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function 
\begin{equation}
	\label{GreensFuncDef} Dh\left( t \right) = \delta \left( t \right), 
\end{equation}
where $D=\frac{d}{dt} + \zeta$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{eqnarray}
	\label{FinalFormContinuous} \frac{dv\left( \mathbf{r},t \right)}{dt} + \zeta v\left( \mathbf{r},t \right) = \\
	\int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)dr'} + u\left(\mathbf{r},t\right). 
\end{eqnarray}
To arrive at the integro-difference equation (IDE) form of the model we discretize time using a first-order Euler method (see~\ref{Time Discretization}) giving 
\begin{eqnarray}
	\label{DiscreteTimeModel} v_{t+1}\left(\mathbf{r}\right) = \xi v_t\left(\mathbf{r}\right) + T_s \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} \\
	+ T_s u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right), 
\end{eqnarray}
where $T_s$ is the time step, $\xi = 1-\zeta T_s$ and $e_t\left(\mathbf{r}\right)$ is an iid spatially coloured, temporally white disturbance term with the covariance 
\begin{equation}
	cov(\mathbf{e}_t,\mathbf{e}_{t+\tau}) = \left\{ 
	\begin{array}{*{20}
		{c}} \gamma\left(\mathbf{r}-\mathbf{r'}\right), & \tau=0 \\
		0 & \mathrm{otherwise}. \\
	\end{array}
	\right. 
\end{equation}
This term was added to drive the system and account for uncertainty in the model. To generate iEEG/LFP data we used the output function 
\begin{equation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)d\mathbf{r}'} + \boldsymbol{\varepsilon}_t, 
\end{equation}
where $\mathbf{r}_n$ defines the location of the sensors in the field where $n=1,...,N$ indexes the sensors and $\boldsymbol{\varepsilon}_t \sim \mathcal{N}\left(0,\Sigma{\varepsilon}\right)$. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the geometry of pick-up range of sensors where 
\begin{equation}
	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)}. 
\end{equation}

@@

	\label{reduced continuous model} \boldsymbol{\phi}^{\top}\left( \mathbf{r} \right) \mathbf{x}_{t+1} = T_s\int_\Omega { f\left( \boldsymbol{\phi} ^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t \right)\boldsymbol{\psi} \left(\mathbf{r}-\mathbf{r}'\right)^{\top}d\mathbf{r}'}\boldsymbol{\theta} \\
	- \xi\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right)x_t + T_s u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right). 

@@

	\label{Homogeneous SS Model} \mathbf{x}_{t+1} &=& T_sq(\mathbf{x}_t)\boldsymbol{\theta} - \xi\mathbf{x}_t + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	&=& T_sQ\left(\mathbf{x}_t,u_t(\mathbf{r})\right) + \mathbf{e}_t \\

@@

% \psfragfig{./Graph/FreqAnalysis}
